Motorcycle Acceleration example

### 2018-10-09 analysis/motorcycle.Rmd
Some more improvements to the motorcycle example.

An illustration of "smoothing via adaptive shrinkage" applied to the
Motorcycle Accident data. This implements the illustration application
described in Sec. 5.1 of the manuscript.

@@

```{r load-pkgs}
library(MASS)
library(lattice)
library(smashr)
library(scales)
library(wavethresh)
library(EbayesThresh)
source("../code/motorcycle.functions.R")
```

@@

Load the motorcycle data, and order the data points by time.

```{r load-data}
data(mcycle)
x.ini.mc <- sort(mcycle$times)
y.ini.mc <- mcycle$accel[order(mcycle$times)]
```

@@

Apply smash to the Motorcycle Accident data set.

@@

```{r plot-smash-estimates}
plot(res.mc$x,res.mc$mu.est,type = "l",
     ylim = c(min(res.mc$y - 2 * sqrt(res.mc$var.est)),
              max(res.mc$y + 2 * sqrt(res.mc$var.est))),
     xlab = "time (ms)", ylab = "acceleration (g)",lwd = 2,
	 col = "darkorange")
lines(res.mc$x, res.mc$mu.est + 2*sqrt(res.mc$var.est),lty = 5,
	 lwd = 2,col = "dodgerblue")
lines(res.mc$x,res.mc$mu.est - 2*sqrt(res.mc$var.est),
      lty = 5,lwd = 2,col = "dodgerblue")
points(res.mc$x,res.mc$y,pch = 1,cex = 1,col = "black")
```

# PREPARE DATA FOR SMASH WITH HOMOGENEOUS VARIANCES
# -------------------------------------------------
# Load the motorcycle data for Empirical Bayes thresholding.
cat("Preparing Motorcycle data for Empirical Bayes thresholding.\n")
x.ini.ti.cons.mc = sort(mcycle$times)
y.ini.ti.cons.mc = mcycle$accel[order(mcycle$times)]

# Load the motorcycle data once more for the second smash analysis.
cat("Preparing Motorcycle data for smash analysis (with equal variances).\n")
x.ini.cons.mc = sort(mcycle$times)
y.ini.cons.mc = mcycle$accel[order(mcycle$times)]

# Apply Empirical Bayes thresholding to the Motorcycle Accident data set.
cat("Running Empirical Bayes thresholding on Motorcycle Accident data set.\n")
res.ti.cons.mc = tithresh.cons.wrapper(x.ini.mc, y.ini.mc)

# Apply smash (with equal variances) to the Motorcycle Accident data set.
cat("Running smash (with equal variances) on Motorcycle Accident data set.\n")
res.cons.mc = smash.cons.wrapper(x.ini.mc, y.ini.mc)

# PREPARE DATA FOR EMPIRICAL BAYES WITH SMASH-ESTIMATED VARIANCES
# ---------------------------------------------------------------
# Load the motorcycle data for Empirical Bayes thresholding (with
# smash-estimated variances).
cat("Preparing data for EbayesThresh (with smash-estimated variances).\n")
x.ini.ti.mc = sort(mcycle$times)
y.ini.ti.mc = mcycle$accel[order(mcycle$times)]

# Apply Empirical Bayes thresholding to the Motorcycle Accident data,
# with variances estimated by smash.
cat(paste("Running EbayesThresh (with smash-estimated variances) on",
          "Motorcycle data.\n"))
res.ti.mc = tithresh.wrapper(x.ini.mc, y.ini.mc)

# SUMMARIZE RESULTS FROM ALL METHODS
# ----------------------------------
# Create another plot showing the Motorcycle Accident data and the
# estimates provided by the four methods.
par(cex.axis = 1.5, cex.sub = 1.8, cex.lab = 1.5,
    mar = c(5.1, 5.1, 2.1, 2.1), mgp = c(3, 1.3, 0))
plot(res.mc$x, res.mc$mu.est, type = 'l',
     ylim = c(min(res.mc$y - 2 * sqrt(res.mc$var.est)),
              max(res.mc$y + 2 * sqrt(res.mc$var.est))),
     xlab = "time (ms)", ylab = "acceleration (g)", lwd = 2)
lines(res.mc$x, res.mc$mu.est + 2 * sqrt(res.mc$var.est), col = 2,
      lty = 5, lwd = 1.8)
lines(res.mc$x, res.mc$mu.est - 2 * sqrt(res.mc$var.est), col = 2,
      lty = 5, lwd = 1.8)
lines(res.ti.cons.mc$x, res.ti.cons.mc$mu.est, lwd = 2, col = 4)
lines(res.cons.mc$x, res.cons.mc$mu.est, lwd = 2, col = 'brown')
lines(res.ti.mc$x, res.ti.mc$mu.est, lwd = 2, col = 'darkgreen')
points(res.mc$x, res.mc$y, cex = 0.7, pch = 16, col = alpha("black", 0.5))
legend('bottomright',
       legend = c('SMASH', 'SMASH, homo', 'TI-thresh, SMASH est',
                  'TI-EBthresh, homo'),
       lty = c(1, 1, 1, 1), lwd = c(2, 2, 2, 2),
       col = c(1, 'brown', 'darkgreen', 4))

>>>>

An illustration of "smoothing via adaptive shrinkage" (SMASH) applied to the
Motorcycle Accident data. This implements the "illustrative application"
presented in Sec. 5.1 of the manuscript.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

@@

```{r load-pkgs, message=FALSE}
library(MASS)
library(lattice)
library(smashr)
library(scales)
library(wavethresh)
library(EbayesThresh)
source("../code/motorcycle.functions.R")
```

@@

Load the motorcycle data from the MASS package, and order the data
points by time.

```{r prepare-data}
data(mcycle)
x.ini.mc <- sort(mcycle$times)
y.ini.mc <- mcycle$accel[order(mcycle$times)]
```

@@

Apply SMASH to the Motorcycle Accident data set.

@@

```{r plot-smash-estimates, fig.height=5, fig.width=7}
plot(res.mc$x,res.mc$mu.est,type = "l",
     ylim = c(min(res.mc$y - 2 * sqrt(res.mc$var.est)),
              max(res.mc$y + 2 * sqrt(res.mc$var.est))),
     xlab = "time (ms)", ylab = "acceleration (g)",lwd = 2,
	 col = "darkorange",xlim = c(0,60),xaxp = c(0,60,6))
lines(res.mc$x, res.mc$mu.est + 2*sqrt(res.mc$var.est),lty = 5,
	 lwd = 2,col = "dodgerblue")
lines(res.mc$x,res.mc$mu.est - 2*sqrt(res.mc$var.est),
      lty = 5,lwd = 2,col = "dodgerblue")
points(res.mc$x,res.mc$y,pch = 1,cex = 1,col = "black")
```

Apply SMASH again, this time assuming equal variances
-----------------------------------------------------

Apply SMASH (assuming with equal variances) to the Motorcycle Accident
data set.

```{r run-smash-homo}
res.cons.mc <- smash.cons.wrapper(x.ini.mc,y.ini.mc)
```

Apply other methods to the Motorcycle Accident data
---------------------------------------------------

Apply Empirical Bayes thresholding to the Motorcycle Accident data set.

```{r run-tithresh}
res.ti.cons.mc <- tithresh.cons.wrapper(x.ini.mc, y.ini.mc)
```

Apply Empirical Bayes thresholding to the Motorcycle Accident data,
this time using the variances estimated by SMASH.

```{r run-tithresh-with-smash-var}
res.ti.mc <- tithresh.wrapper(x.ini.mc,y.ini.mc)
```

SUMMARIZE RESULTS FROM ALL METHODS
----------------------------------

Create another plot showing the Motorcycle Accident data and the
estimates provided by the four methods.

```{r}
plot(res.mc$x, res.mc$mu.est, type = 'l',
     ylim = c(min(res.mc$y - 2 * sqrt(res.mc$var.est)),
              max(res.mc$y + 2 * sqrt(res.mc$var.est))),
     xlab = "time (ms)", ylab = "acceleration (g)", lwd = 2)
lines(res.mc$x, res.mc$mu.est + 2 * sqrt(res.mc$var.est), col = 2,
      lty = 5, lwd = 1.8)
lines(res.mc$x, res.mc$mu.est - 2 * sqrt(res.mc$var.est), col = 2,
      lty = 5, lwd = 1.8)
lines(res.ti.cons.mc$x, res.ti.cons.mc$mu.est, lwd = 2, col = 4)
lines(res.cons.mc$x, res.cons.mc$mu.est, lwd = 2, col = 'brown')
lines(res.ti.mc$x, res.ti.mc$mu.est, lwd = 2, col = 'darkgreen')
points(res.mc$x, res.mc$y, cex = 0.7, pch = 16, col = alpha("black", 0.5))
legend('bottomright',
       legend = c('SMASH', 'SMASH, homo', 'TI-thresh, SMASH est',
                  'TI-EBthresh, homo'),
       lty = c(1, 1, 1, 1), lwd = c(2, 2, 2, 2),
       col = c(1, 'brown', 'darkgreen', 4))
```
